---
title: "Getting Started with datadriftR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with datadriftR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

## What is Data Drift?

Data drift occurs when the statistical properties of data change over time. This is critical in machine learning systems where models may degrade as data patterns shift.

## Available Methods

**datadriftR** provides 8 drift detection methods:

| Method | Best For | Key Advantage |
|:-------|:---------|:--------------|
| [DDM](../reference/DDM.html) | Binary error streams | Simple and fast |
| [EDDM](../reference/EDDM.html) | Binary error streams | Earlier detection than DDM |
| [HDDM-A](../reference/HDDM_A.html) | Binary streams | Theoretical guarantees |
| [HDDM-W](../reference/HDDM_W.html) | Binary streams | Memory efficient |
| [KSWIN](../reference/KSWIN.html) | Any numeric data | Distribution comparison |
| [Page-Hinkley](../reference/PageHinkley.html) | Continuous data | Detects mean changes |
| [KL-Divergence](../reference/KLDivergence.html) | Continuous data | Information-theoretic |
| [Profile Difference](../reference/ProfileDifference.html) | PDP comparison | Detects model behavior changes |

## Quick Start: Using detect_drift()

The simplest way to detect drift - process entire stream at once:

```{r simple-example}
library(datadriftR)
set.seed(123)

n_stable <- 500
n_drift <- 500

stream <- c(
  sample(c(0, 1), n_stable, replace = TRUE, prob = c(0.7, 0.3)),
  sample(c(0, 1), n_drift, replace = TRUE, prob = c(0.3, 0.7))
)

drift_point <- n_stable + 1

results <- detect_drift(stream, method = "ddm", include_warnings = FALSE)
head(results, 3)
```

## All Methods Comparison

```{r all-methods}
methods <- c("ddm", "eddm", "hddm_a", "hddm_w", "kswin", "page_hinkley")

comparison <- do.call(rbind, lapply(methods, function(m) {
  res <- detect_drift(stream, method = m, include_warnings = FALSE)
  data.frame(
    Method = toupper(m),
    Detections = nrow(res),
    First = if (nrow(res) > 0) min(res$index) else NA
  )
}))

comparison
```

```{r comparison-plot, fig.cap="Detection points by method", fig.height=5}
roll_mean <- sapply(seq_along(stream), function(i) mean(stream[max(1,i-50):i]))

plot(roll_mean, type = "l", col = "gray40", lwd = 2,
     xlab = "Observation", ylab = "Rolling Mean",
     main = "Drift Detection Comparison")
abline(v = drift_point, col = "black", lty = 2, lwd = 2)

colors <- c("#E63946", "#F4A261", "#2A9D8F", "#9B5DE5", "#00BBF9", "#8B4513")
for (i in seq_len(nrow(comparison))) {
  if (!is.na(comparison$First[i])) {
    abline(v = comparison$First[i], col = colors[i], lwd = 2)
  }
}

legend("bottomleft",
       legend = c(paste0("Actual (", drift_point, ")"),
                  paste0(comparison$Method, " (", comparison$First, ")")),
       col = c("black", colors),
       lty = c(2, rep(1, 6)), lwd = 2, cex = 0.8, bg = "white")
```

## Online Learning: Streaming with For Loop

For real-time applications, process observations one by one:

```{r online-ddm}
ddm <- DDM$new()

for (i in seq_along(stream)) {
  ddm$add_element(stream[i])
  if (ddm$change_detected) {
    cat("DDM detected drift at index:", i, "\n")
    ddm$reset()
  }
}
```

```{r online-kswin}
kswin <- KSWIN$new(alpha = 0.005, window_size = 100)

for (i in seq_along(stream)) {
  kswin$add_element(stream[i])
  if (kswin$detected_change()) {
    cat("KSWIN detected drift at index:", i, "\n")
  }
}
```

```{r online-pagehinkley}
ph <- PageHinkley$new(threshold = 50)

for (i in seq_along(stream)) {
  ph$add_element(stream[i])
  if (ph$detected_change()) {
    cat("Page-Hinkley detected drift at index:", i, "\n")
    ph$reset()
  }
}
```

## Continuous Data Example

```{r continuous, fig.cap="Variance change detection in continuous data"}
set.seed(111)

n_before <- 200
n_after <- 200

before <- rnorm(n_before, mean = 0, sd = 1)
after <- rnorm(n_after, mean = 0, sd = 3)
cont_stream <- c(before, after)
cont_drift <- n_before + 1

results_cont <- detect_drift(cont_stream, method = "kswin",
                              alpha = 0.001, window_size = 50, stat_size = 25)

plot(cont_stream, type = "l", col = "gray50",
     xlab = "Observation", ylab = "Value", main = "Variance Change Detection")
abline(v = cont_drift, col = "black", lty = 2, lwd = 2)
if (nrow(results_cont) > 0) {
  abline(v = results_cont$index[1], col = "red", lwd = 2)
  legend("topleft",
         c(paste0("Actual (", cont_drift, ")"),
           paste0("Detected (", results_cont$index[1], ")")),
         col = c("black", "red"), lty = c(2, 1), lwd = 2)
}
```

## Summary

| Approach | Function | Use Case |
|----------|----------|----------|
| Batch | `detect_drift()` | Analyze complete dataset |
| Online | `DDM$new()`, `KSWIN$new()`, etc. | Real-time streaming |

For more details: `?detect_drift`, `?DDM`, `?KSWIN`, `?PageHinkley`
