---
title: "Getting Started with datadriftR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with datadriftR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## What is Concept Drift?

In machine learning systems, **concept drift** occurs when the statistical properties of the data change over time. This can cause model performance to degrade because the patterns learned during training no longer apply to new data.

For example:
- A spam filter trained on old emails may fail to catch new spam techniques
- A fraud detection model may miss new types of fraudulent behavior
- A recommendation system may become less accurate as user preferences evolve

**datadriftR** helps you detect these changes in real-time by monitoring your data streams and alerting you when drift occurs.

## Quick Start

The `detect_drift()` function is the easiest way to detect drift in your data:

```{r quickstart}
library(datadriftR)
set.seed(123)

# Create a data stream with a sudden change at position 501
# First 500 observations: 30% ones, 70% zeros
# Last 500 observations: 70% ones, 30% zeros (drift!)
stable_period <- sample(c(0, 1), 500, replace = TRUE, prob = c(0.7, 0.3))
drift_period  <- sample(c(0, 1), 500, replace = TRUE, prob = c(0.3, 0.7))
stream <- c(stable_period, drift_period)

# Detect drift using DDM method
results <- detect_drift(stream, method = "ddm")
print(results)
```

The function returns a data frame with all detected drift points and their positions in the stream.

## Available Methods

datadriftR provides 8 different drift detection methods. Each method has its own strengths:

| Method | Best For | Description |
|--------|----------|-------------|
| `ddm` | Binary streams | Monitors error rate changes |
| `eddm` | Binary streams | Detects drift earlier than DDM |
| `hddm_a` | Binary streams | Uses Hoeffding bounds (averaging) |
| `hddm_w` | Binary streams | Uses Hoeffding bounds (weighted) |
| `kswin` | Any numeric data | Kolmogorov-Smirnov statistical test |
| `page_hinkley` | Any numeric data | Cumulative sum approach |
| `kl_divergence` | Any numeric data | Information-theoretic approach |
| `profile_difference` | Functional data | Compares data profiles |

## Comparing Methods

Let's see how different methods perform on the same data:

```{r comparison}
methods <- c("ddm", "eddm", "hddm_a", "page_hinkley", "kswin")

results_list <- lapply(methods, function(m) {
  res <- detect_drift(stream, method = m, include_warnings = FALSE)
  data.frame(
    Method = toupper(m),
    Detections = nrow(res),
    First_Detection = if(nrow(res) > 0) min(res$index) else NA
  )
})

comparison <- do.call(rbind, results_list)
print(comparison)
```

Since the actual drift happens at position 501, methods detecting drift closer to this point are more accurate.

## Understanding Warnings and Drift

Some methods (DDM, EDDM, HDDM) provide two levels of alerts:

- **Warning**: Early signal that drift might be occurring
- **Drift**: Confirmed detection of concept drift

```{r warnings}
# Get both warnings and drift detections
all_alerts <- detect_drift(stream, method = "ddm", include_warnings = TRUE)
table(all_alerts$type)

# Get only confirmed drift
drift_only <- detect_drift(stream, method = "ddm", include_warnings = FALSE)
nrow(drift_only)
```

## Tuning Detection Sensitivity

Each method has parameters you can adjust to control sensitivity:

```{r tuning}
# KSWIN: Lower alpha = more conservative (fewer false alarms)
conservative <- detect_drift(stream, method = "kswin",
                             alpha = 0.001, window_size = 100)

# KSWIN: Higher alpha = more sensitive (catches drift faster)
sensitive <- detect_drift(stream, method = "kswin",
                          alpha = 0.01, window_size = 100)

cat("Conservative approach:", nrow(conservative), "detections\n")
cat("Sensitive approach:", nrow(sensitive), "detections\n")
```

## Working with Continuous Data

KSWIN and Page-Hinkley work well with continuous numeric data:

```{r continuous}
set.seed(42)

# Continuous stream: mean shifts from 0 to 1 at position 301
continuous_stream <- c(
  rnorm(300, mean = 0, sd = 1),
  rnorm(300, mean = 1, sd = 1)
)

# Detect the mean shift
results <- detect_drift(continuous_stream, method = "kswin",
                        alpha = 0.005, window_size = 100, stat_size = 30)

if(nrow(results) > 0) {
  cat("Drift detected at position:", min(results$index), "\n")
  cat("(Actual drift occurred at position 301)\n")
}
```

## Real-World Example: Monitoring Model Errors

Here's how you might use datadriftR in practice to monitor a model's prediction errors:

```{r realworld}
set.seed(123)

# Simulate model predictions over time
# Initially the model works well (low error rate)
# Later, drift causes more errors
good_predictions <- sample(c(0, 1), 400, replace = TRUE, prob = c(0.9, 0.1))
bad_predictions  <- sample(c(0, 1), 200, replace = TRUE, prob = c(0.6, 0.4))
error_stream <- c(good_predictions, bad_predictions)

# Monitor for performance degradation
alerts <- detect_drift(error_stream, method = "ddm", include_warnings = FALSE)

if(nrow(alerts) > 0) {
  cat("Model performance degradation detected!\n")
  cat("First alert at observation:", min(alerts$index), "\n")
  cat("Action: Consider retraining the model with recent data.\n")
}
```

## Choosing the Right Method

**For binary/error streams (0s and 1s):**
- Start with `ddm` - it's simple and effective
- Try `eddm` if you need earlier detection
- Use `hddm_a` or `hddm_w` for theoretical guarantees

**For continuous numeric data:**
- `kswin` is versatile and statistically grounded
- `page_hinkley` works well for detecting mean changes

**General tips:**
- Test multiple methods on your specific data
- Adjust parameters based on your tolerance for false alarms vs. missed detections
- Consider using warnings as early indicators before confirmed drift

## Learn More

- See `?detect_drift` for full parameter documentation
- Check individual method documentation: `?DDM`, `?KSWIN`, `?PageHinkley`
- Visit the [GitHub repository](https://github.com/ugurdar/datadriftR) for more examples
