---
title: "Comprehensive Guide to Drift Detection with datadriftR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comprehensive Guide to Drift Detection with datadriftR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

Machine learning models are trained on historical data with the assumption that future data will follow similar patterns. However, in real-world applications, **data evolves over time**. When the statistical properties of data change, model performance can silently degrade—leading to incorrect predictions, financial losses, or even safety risks.

**datadriftR** provides a comprehensive toolkit for detecting these changes in streaming data, enabling proactive model maintenance and reliable production ML systems.

# Understanding Drift: A Conceptual Framework

## What is Dataset Shift?

**Dataset shift** is the umbrella term for any scenario where the joint distribution of inputs and outputs differs between training and deployment:

$$P_{train}(X, Y) \neq P_{test}(X, Y)$$

This shift can manifest in several distinct forms, each requiring different detection strategies.

## Types of Drift

### 1. Covariate Shift (Data Drift / Virtual Drift)

The input distribution changes while the relationship between inputs and outputs remains the same:

$$P_{train}(X) \neq P_{test}(X) \quad \text{but} \quad P(Y|X) \text{ unchanged}$$

**Real-world examples:**

- **E-commerce**: Customer demographics shift from urban millennials to suburban families
- **Healthcare**: Hospital starts serving older patient population
- **Finance**: Economic recession changes spending patterns

```{r covariate-shift-viz, fig.cap="Covariate Shift: Input distribution changes", fig.height=4}
set.seed(42)
par(mfrow = c(1, 2))

# Training data
x_train <- rnorm(500, mean = 2, sd = 1)
hist(x_train, breaks = 30, col = "#2A9D8F", border = "white",
     main = "Training Distribution", xlab = "Feature X", xlim = c(-2, 10))

# Test data (shifted)
x_test <- rnorm(500, mean = 5, sd = 1.5)
hist(x_test, breaks = 30, col = "#E63946", border = "white",
     main = "Production Distribution", xlab = "Feature X", xlim = c(-2, 10))

par(mfrow = c(1, 1))
```

### 2. Concept Drift (Real Drift)

The relationship between inputs and outputs changes:

$$P_{train}(Y|X) \neq P_{test}(Y|X)$$

**Real-world examples:**

- **Spam detection**: Spammers adapt strategies to evade filters
- **Fraud detection**: Criminals develop new fraud techniques
- **Recommendation systems**: User preferences evolve (e.g., pandemic changes movie preferences)

```{r concept-drift-viz, fig.cap="Concept Drift: Decision boundary shifts", fig.height=4}
set.seed(123)
par(mfrow = c(1, 2))

# Before concept drift
x <- seq(0, 10, length.out = 100)
y_before <- 0.5 * x + rnorm(100, sd = 0.5)
plot(x, y_before, pch = 19, col = "#2A9D8F", cex = 0.7,
     main = "Before Concept Drift", xlab = "X", ylab = "Y")
abline(lm(y_before ~ x), col = "#1D3557", lwd = 2)

# After concept drift (relationship changed)
y_after <- -0.3 * x + 8 + rnorm(100, sd = 0.5)
plot(x, y_after, pch = 19, col = "#E63946", cex = 0.7,
     main = "After Concept Drift", xlab = "X", ylab = "Y")
abline(lm(y_after ~ x), col = "#1D3557", lwd = 2)

par(mfrow = c(1, 1))
```

### 3. Prior Probability Shift (Label Drift)

The class distribution changes:

$$P_{train}(Y) \neq P_{test}(Y)$$

**Real-world examples:**

- **Manufacturing**: Defect rate increases due to equipment wear
- **Medical diagnosis**: Disease prevalence changes seasonally
- **Churn prediction**: Economic downturn increases churn rates

### Drift Patterns Over Time

Drift doesn't always happen abruptly. Understanding the **temporal pattern** helps choose the right detector:

```{r drift-patterns, fig.cap="Different drift patterns over time", fig.height=5}
set.seed(456)
n <- 500
t <- 1:n

par(mfrow = c(2, 2))

# 1. Abrupt drift
abrupt <- c(rnorm(250, 0, 1), rnorm(250, 3, 1))
plot(t, abrupt, type = "l", col = "#457B9D", lwd = 1.5,
     main = "Abrupt Drift", xlab = "Time", ylab = "Value")
abline(v = 250, col = "#E63946", lty = 2, lwd = 2)

# 2. Gradual drift
gradual <- sapply(t, function(i) {
  p <- min(1, max(0, (i - 200) / 100))
  rnorm(1, mean = p * 3, sd = 1)
})
plot(t, gradual, type = "l", col = "#457B9D", lwd = 1.5,
     main = "Gradual Drift", xlab = "Time", ylab = "Value")
rect(200, -5, 300, 8, col = rgb(0.9, 0.3, 0.3, 0.2), border = NA)

# 3. Incremental drift
incremental <- rnorm(n, mean = t/100, sd = 1)
plot(t, incremental, type = "l", col = "#457B9D", lwd = 1.5,
     main = "Incremental Drift", xlab = "Time", ylab = "Value")

# 4. Recurring drift (seasonal)
recurring <- sin(t / 30) * 2 + rnorm(n, 0, 0.5)
plot(t, recurring, type = "l", col = "#457B9D", lwd = 1.5,
     main = "Recurring/Seasonal Drift", xlab = "Time", ylab = "Value")

par(mfrow = c(1, 1))
```

## Why Drift Detection Matters

| Consequence | Description | Impact |
|:------------|:------------|:-------|
| **Silent failures** | Model degrades without alerts | Wrong decisions accumulate |
| **Delayed response** | By the time accuracy drops, damage is done | Revenue loss, safety risks |
| **Wasted resources** | Unnecessary retraining when no drift | Computational costs |
| **Regulatory issues** | Models must be explainable and monitored | Compliance failures |

# datadriftR: Your Drift Detection Toolkit

## Available Methods

**datadriftR** provides 9 complementary drift detection methods, each suited for different scenarios:

### Error-Rate Based Methods (Supervised)

These methods monitor **prediction errors** and detect when error patterns change:

| Method | Description | Best For |
|:-------|:------------|:---------|
| [DDM](../reference/DDM.html) | Drift Detection Method | Fast detection of abrupt changes |
| [EDDM](../reference/EDDM.html) | Early DDM | Gradual drift, earlier warnings |
| [HDDM-A](../reference/HDDM_A.html) | Hoeffding with averaging | Theoretical guarantees |
| [HDDM-W](../reference/HDDM_W.html) | Hoeffding with weighting | Memory efficient |

### Distribution-Based Methods (Unsupervised)

These methods compare **data distributions** directly:

| Method | Description | Best For |
|:-------|:------------|:---------|
| [KSWIN](../reference/KSWIN.html) | Kolmogorov-Smirnov Windowing | Any numeric data, robust |
| [ADWIN](../reference/ADWIN.html) | Adaptive Windowing | Automatic window sizing |
| [Page-Hinkley](../reference/PageHinkley.html) | Cumulative sum test | Mean shift detection |
| [KL-Divergence](../reference/KLDivergence.html) | Information-theoretic | Histogram comparison |

### Model-Based Methods

| Method | Description | Best For |
|:-------|:------------|:---------|
| [ProfileDifference](../reference/ProfileDifference.html) | PDP comparison | Model behavior changes |

## How to Choose a Method?

```
Do you have ground truth labels?
├── YES → Use error-rate methods (DDM, EDDM, HDDM)
│         ├── Need fast detection? → DDM
│         ├── Gradual drift? → EDDM
│         └── Need guarantees? → HDDM-A/W
│
└── NO → Use distribution methods
          ├── Continuous data? → KSWIN, Page-Hinkley, ADWIN
          ├── Categorical? → KL-Divergence
          └── Model behavior? → ProfileDifference
```

# Quick Start: Basic Usage

## The `detect_drift()` Function

The simplest way to detect drift—process an entire stream at once:

The simplest way to detect drift - process entire stream at once:

```{r simple-example}
library(datadriftR)
set.seed(123)

n_stable <- 500
n_drift <- 500

stream <- c(
  sample(c(0, 1), n_stable, replace = TRUE, prob = c(0.7, 0.3)),
  sample(c(0, 1), n_drift, replace = TRUE, prob = c(0.3, 0.7))
)

drift_point <- n_stable + 1

results <- detect_drift(stream, method = "ddm", include_warnings = FALSE)
head(results, 3)
```

## All Methods Comparison

```{r all-methods}
methods <- c("ddm", "eddm", "hddm_a", "hddm_w", "kswin", "adwin", "page_hinkley")

comparison <- do.call(rbind, lapply(methods, function(m) {
  res <- detect_drift(stream, method = m, include_warnings = FALSE)
  data.frame(
    Method = toupper(m),
    Detections = nrow(res),
    First = if (nrow(res) > 0) min(res$index) else NA
  )
}))

comparison
```

```{r comparison-plot, fig.cap="Detection points by method", fig.height=5}
roll_mean <- sapply(seq_along(stream), function(i) mean(stream[max(1,i-50):i]))

plot(roll_mean, type = "l", col = "gray40", lwd = 2,
     xlab = "Observation", ylab = "Rolling Mean",
     main = "Drift Detection Comparison")
abline(v = drift_point, col = "black", lty = 2, lwd = 2)

colors <- c("#E63946", "#F4A261", "#2A9D8F", "#9B5DE5", "#00BBF9", "#8B4513", "#FF69B4")
for (i in seq_len(nrow(comparison))) {
  if (!is.na(comparison$First[i])) {
    abline(v = comparison$First[i], col = colors[i], lwd = 2)
  }
}

legend("bottomleft",
       legend = c(paste0("Actual (", drift_point, ")"),
                  paste0(comparison$Method, " (", comparison$First, ")")),
       col = c("black", colors),
       lty = c(2, rep(1, 6)), lwd = 2, cex = 0.8, bg = "white")
```

## Online Learning: Streaming with For Loop

For real-time applications, process observations one by one:

```{r online-ddm}
ddm <- DDM$new()

for (i in seq_along(stream)) {
  ddm$add_element(stream[i])
  if (ddm$change_detected) {
    cat("DDM detected drift at index:", i, "\n")
    ddm$reset()
  }
}
```

```{r online-kswin}
kswin <- KSWIN$new(alpha = 0.005, window_size = 100)

for (i in seq_along(stream)) {
  kswin$add_element(stream[i])
  if (kswin$detected_change()) {
    cat("KSWIN detected drift at index:", i, "\n")
  }
}
```

```{r online-pagehinkley}
ph <- PageHinkley$new(threshold = 50)

for (i in seq_along(stream)) {
  ph$add_element(stream[i])
  if (ph$detected_change()) {
    cat("Page-Hinkley detected drift at index:", i, "\n")
    ph$reset()
  }
}
```

```{r online-adwin}
adwin <- ADWIN$new(delta = 0.002)

for (i in seq_along(stream)) {
  adwin$add_element(stream[i])
  if (adwin$detected_change()) {
    cat("ADWIN detected drift at index:", i, "\n")
  }
}
```

## Continuous Data Example

```{r continuous, fig.cap="Variance change detection in continuous data"}
set.seed(111)

n_before <- 200
n_after <- 200

before <- rnorm(n_before, mean = 0, sd = 1)
after <- rnorm(n_after, mean = 0, sd = 3)
cont_stream <- c(before, after)
cont_drift <- n_before + 1

results_cont <- detect_drift(cont_stream, method = "kswin",
                              alpha = 0.001, window_size = 50, stat_size = 25)

plot(cont_stream, type = "l", col = "gray50",
     xlab = "Observation", ylab = "Value", main = "Variance Change Detection")
abline(v = cont_drift, col = "black", lty = 2, lwd = 2)
if (nrow(results_cont) > 0) {
  abline(v = results_cont$index[1], col = "red", lwd = 2)
  legend("topleft",
         c(paste0("Actual (", cont_drift, ")"),
           paste0("Detected (", results_cont$index[1], ")")),
         col = c("black", "red"), lty = c(2, 1), lwd = 2)
}
```

## Summary

| Approach | Function | Use Case |
|----------|----------|----------|
| Batch | `detect_drift()` | Analyze complete dataset |
| Online | `DDM$new()`, `KSWIN$new()`, etc. | Real-time streaming |

For more details: `?detect_drift`, `?DDM`, `?KSWIN`, `?PageHinkley`

# Real-World Scenarios

## Scenario 1: Monitoring Model Errors (Classification)

You have a deployed classification model and want to detect when it starts making more errors:

```{r scenario-classification}
set.seed(789)

# Simulate prediction errors: 0 = correct, 1 = error
# Model works well initially (5% error), then degrades (25% error)
errors <- c(
  rbinom(400, 1, prob = 0.05),  # Good performance
  rbinom(600, 1, prob = 0.25)   # Degraded performance
)

# Detect when model starts failing
drift_results <- detect_drift(errors, method = "ddm", include_warnings = TRUE)
print(drift_results)
```

## Scenario 2: Sensor Data Monitoring (IoT)

A temperature sensor should maintain consistent readings, but equipment failure causes a shift:

```{r scenario-sensor, fig.cap="Sensor drift detection"}
set.seed(321)

# Normal operation: 20°C with some noise
normal_readings <- rnorm(300, mean = 20, sd = 1)

# Equipment malfunction: readings drift upward
faulty_readings <- rnorm(200, mean = 28, sd = 2)

sensor_stream <- c(normal_readings, faulty_readings)

# Use Page-Hinkley for continuous data
ph <- PageHinkley$new(threshold = 20, delta = 0.01)
drift_points <- c()

for (i in seq_along(sensor_stream)) {
  ph$add_element(sensor_stream[i])
  if (ph$detected_change()) {
    drift_points <- c(drift_points, i)
    ph$reset()
  }
}

# Visualize
plot(sensor_stream, type = "l", col = "#457B9D",
     xlab = "Time", ylab = "Temperature (°C)",
     main = "Sensor Drift Detection")
abline(v = 300, col = "black", lty = 2, lwd = 2)
if (length(drift_points) > 0) {
  abline(v = drift_points[1], col = "#E63946", lwd = 2)
  legend("topleft", c("Actual malfunction (300)", paste0("Detected (", drift_points[1], ")")),
         col = c("black", "#E63946"), lty = c(2, 1), lwd = 2)
}
```

## Scenario 3: Gradual Drift Detection

Some drift happens slowly—like user behavior changing over months:

```{r scenario-gradual, fig.cap="Detecting gradual drift with ADWIN"}
set.seed(555)

# Gradual shift in user engagement
n <- 1000
engagement <- numeric(n)
for (i in 1:n) {
  # Engagement slowly decreases over time
  base_rate <- 0.6 - (i / n) * 0.4  # From 60% to 20%
  engagement[i] <- rbinom(1, 1, prob = base_rate)
}

# ADWIN adapts its window to detect gradual changes
adwin <- ADWIN$new(delta = 0.001)
drift_points <- c()

for (i in seq_along(engagement)) {
  adwin$add_element(engagement[i])
  if (adwin$detected_change()) {
    drift_points <- c(drift_points, i)
  }
}

cat("ADWIN detected gradual drift at indices:", drift_points, "\n")

# Visualize with rolling average
roll_avg <- sapply(seq_along(engagement), function(i) mean(engagement[max(1,i-50):i]))
plot(roll_avg, type = "l", col = "#457B9D", lwd = 2,
     xlab = "Time", ylab = "Rolling Engagement Rate",
     main = "Gradual Drift in User Engagement")
if (length(drift_points) > 0) {
  abline(v = drift_points, col = "#E63946", lty = 2)
}
```

# Deep Dive: Each Method Explained

## DDM (Drift Detection Method)

**How it works**: Monitors error rate and standard deviation. Drift is signaled when:
$$p_i + s_i \geq p_{min} + 3 \cdot s_{min}$$

Where $p_i$ is error rate and $s_i$ is standard deviation at observation $i$.

```{r ddm-deep}
ddm <- DDM$new(
  min_num_instances = 30,  # Minimum observations before checking
  warning_level = 2.0,     # Standard deviations for warning
  out_control_level = 3.0  # Standard deviations for drift
)

# Process stream
for (i in seq_along(stream)) {
  ddm$add_element(stream[i])
  
  if (ddm$warning_detected && !ddm$change_detected) {
    cat("DDM WARNING at index:", i, "\n")
  }
  if (ddm$change_detected) {
    cat("DDM DRIFT at index:", i, "\n")
    ddm$reset()
  }
}
```

## ADWIN (Adaptive Windowing)

**How it works**: Maintains a variable-size window and automatically shrinks it when drift is detected. Uses Hoeffding bounds to determine if two sub-windows have significantly different means.

**Key advantage**: No need to specify window size—it adapts automatically!

```{r adwin-deep}
adwin <- ADWIN$new(
  delta = 0.002,           # Confidence parameter (smaller = more sensitive)
  clock = 32,              # Check frequency
  max_buckets = 5,         # Memory parameter
  min_window_length = 5,   # Minimum window
  grace_period = 10        # Initial warm-up
)

drift_indices <- c()
for (i in seq_along(stream)) {
  adwin$add_element(stream[i])
  if (adwin$detected_change()) {
    drift_indices <- c(drift_indices, i)
  }
}
cat("ADWIN detected drift at:", drift_indices, "\n")
```

## KSWIN (Kolmogorov-Smirnov Windowing)

**How it works**: Compares the distribution of a sliding window against a reference window using the two-sample Kolmogorov-Smirnov test.

```{r kswin-deep}
kswin <- KSWIN$new(
  alpha = 0.005,      # Significance level
  window_size = 100,  # Reference window size
  stat_size = 30      # Sliding window size
)

for (i in seq_along(stream)) {
  kswin$add_element(stream[i])
  if (kswin$detected_change()) {
    cat("KSWIN detected drift at index:", i, "\n")
  }
}
```

## Page-Hinkley Test

**How it works**: Monitors cumulative sum of differences from the mean. Detects when the cumulative deviation exceeds a threshold.

$$PH_t = \sum_{i=1}^{t}(x_i - \bar{x}_t - \delta)$$

Drift is signaled when: $PH_t - \min(PH) > \lambda$

```{r pagehinkley-deep}
ph <- PageHinkley$new(
  delta = 0.005,    # Minimum change magnitude to detect
  threshold = 50,   # Detection threshold (lambda)
  alpha = 0.9999    # Forgetting factor for mean
)

for (i in seq_along(stream)) {
  ph$add_element(stream[i])
  if (ph$detected_change()) {
    cat("Page-Hinkley detected drift at index:", i, "\n")
    ph$reset()
  }
}
```

# Method Comparison on Different Drift Types

## Abrupt vs. Gradual Drift

```{r comparison-drift-types, fig.cap="Method performance on different drift types", fig.height=6}
set.seed(999)

# Create two streams
n <- 800

# Abrupt drift
abrupt_stream <- c(
  rbinom(400, 1, 0.3),
  rbinom(400, 1, 0.7)
)

# Gradual drift  
gradual_stream <- sapply(1:n, function(i) {
  p <- 0.3 + (i > 300) * min(0.4, (i - 300) / 250)
  rbinom(1, 1, p)
})

methods <- c("ddm", "eddm", "adwin", "page_hinkley")

results_abrupt <- lapply(methods, function(m) {
  res <- detect_drift(abrupt_stream, method = m, include_warnings = FALSE)
  if (nrow(res) > 0) min(res$index) else NA
})

results_gradual <- lapply(methods, function(m) {
  res <- detect_drift(gradual_stream, method = m, include_warnings = FALSE)
  if (nrow(res) > 0) min(res$index) else NA
})

comparison_df <- data.frame(
  Method = toupper(methods),
  Abrupt_First = unlist(results_abrupt),
  Gradual_First = unlist(results_gradual)
)

print(comparison_df)

# Visualization
par(mfrow = c(2, 1))

roll_abrupt <- sapply(seq_along(abrupt_stream), function(i) mean(abrupt_stream[max(1,i-30):i]))
plot(roll_abrupt, type = "l", col = "#457B9D", lwd = 2,
     main = "Abrupt Drift Detection", xlab = "Time", ylab = "Rolling Mean")
abline(v = 400, col = "black", lty = 2, lwd = 2)
colors <- c("#E63946", "#F4A261", "#2A9D8F", "#9B5DE5")
for (i in seq_along(methods)) {
  if (!is.na(comparison_df$Abrupt_First[i])) {
    abline(v = comparison_df$Abrupt_First[i], col = colors[i], lwd = 2)
  }
}
legend("bottomleft", c("Actual", comparison_df$Method), 
       col = c("black", colors), lty = c(2, rep(1, 4)), lwd = 2, cex = 0.7)

roll_gradual <- sapply(seq_along(gradual_stream), function(i) mean(gradual_stream[max(1,i-30):i]))
plot(roll_gradual, type = "l", col = "#457B9D", lwd = 2,
     main = "Gradual Drift Detection", xlab = "Time", ylab = "Rolling Mean")
abline(v = 300, col = "black", lty = 2, lwd = 2)
for (i in seq_along(methods)) {
  if (!is.na(comparison_df$Gradual_First[i])) {
    abline(v = comparison_df$Gradual_First[i], col = colors[i], lwd = 2)
  }
}
legend("bottomleft", c("Drift Start", comparison_df$Method), 
       col = c("black", colors), lty = c(2, rep(1, 4)), lwd = 2, cex = 0.7)

par(mfrow = c(1, 1))
```

# Best Practices

## 1. Start with Multiple Detectors

Different methods have different strengths. Run several in parallel:

```{r best-practice-multiple}
run_all_detectors <- function(stream) {
  methods <- c("ddm", "eddm", "hddm_a", "adwin", "kswin", "page_hinkley")
  
  results <- lapply(methods, function(m) {
    res <- detect_drift(stream, method = m, include_warnings = FALSE)
    if (nrow(res) > 0) res$index else integer(0)
  })
  names(results) <- methods
  
  # Consensus: drift confirmed if majority agree
  all_detections <- unlist(results)
  if (length(all_detections) > 0) {
    hist(all_detections, breaks = 20, main = "Detection Distribution",
         xlab = "Index", col = "#2A9D8F")
  }
  
  return(results)
}

detections <- run_all_detectors(stream)
```

## 2. Tune Parameters for Your Domain

```{r best-practice-tuning}
# More sensitive detection (more false positives)
sensitive <- detect_drift(stream, method = "ddm", 
                          out_control_level = 2.5)

# Less sensitive (fewer false positives)
conservative <- detect_drift(stream, method = "ddm", 
                             out_control_level = 3.5)

cat("Sensitive detections:", nrow(sensitive), "\n")
cat("Conservative detections:", nrow(conservative), "\n")
```

## 3. Monitor Multiple Features

In practice, monitor each important feature independently:

```{r best-practice-multivariate}
# Simulate multivariate data
set.seed(123)
feature1 <- c(rnorm(300, 0, 1), rnorm(200, 2, 1))
feature2 <- c(rnorm(300, 5, 2), rnorm(200, 5, 2))  # No drift
feature3 <- c(rnorm(400, 10, 1), rnorm(100, 15, 1))

features <- list(feature1 = feature1, feature2 = feature2, feature3 = feature3)

for (name in names(features)) {
  result <- detect_drift(features[[name]], method = "kswin", 
                         alpha = 0.01, window_size = 50)
  cat(name, ": ", nrow(result), "drift(s) detected\n")
}
```

# Summary Table

| Method | Data Type | Drift Type | Memory | Speed | Use When |
|:-------|:----------|:-----------|:-------|:------|:---------|
| DDM | Binary errors | Abrupt | Low | Fast | Classification monitoring |
| EDDM | Binary errors | Gradual | Low | Fast | Early warning needed |
| HDDM-A | Binary | Both | Medium | Medium | Need guarantees |
| HDDM-W | Binary | Both | Low | Fast | Memory constrained |
| ADWIN | Numeric | Both | Adaptive | Medium | Unknown drift type |
| KSWIN | Numeric | Abrupt | Fixed | Medium | Distribution changes |
| Page-Hinkley | Continuous | Abrupt | Low | Fast | Mean shift |
| KL-Divergence | Categorical | Any | Fixed | Medium | Histogram data |
| ProfileDiff | Model PDPs | Any | High | Slow | Model behavior |

# Further Resources

- **API Reference**: See the [Reference](../reference/index.html) section for detailed function documentation
- **Paper**: Dar & Cavus (2025). datadriftR: An R package for data drift detection. *Journal of Open Source Software*.
- **Python alternatives**: [River](https://riverml.xyz/), [scikit-multiflow](https://scikit-multiflow.github.io/)

For questions or issues, please visit our [GitHub repository](https://github.com/ugurdar/datadriftR).
