---
title: "Getting Started with datadriftR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with datadriftR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

Machine learning models assume that future data follows similar patterns to training data. In production, **data evolves over time**—and when statistical properties change, model performance silently degrades.

**datadriftR** detects these changes in streaming data, enabling proactive model maintenance.

# Types of Drift

```{r drift-types-viz, fig.height=5, fig.cap="Common drift patterns"}
set.seed(42)
n <- 400
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

# Abrupt
plot(c(rnorm(200, 0, 1), rnorm(200, 3, 1)), type = "l", col = "#457B9D",
     main = "Abrupt Drift", xlab = "Time", ylab = "Value")
abline(v = 200, col = "#E63946", lty = 2, lwd = 2)

# Gradual
gradual <- sapply(1:n, function(i) rnorm(1, mean = min(3, max(0, (i-150)/80)), sd = 1))
plot(gradual, type = "l", col = "#457B9D",
     main = "Gradual Drift", xlab = "Time", ylab = "Value")
rect(150, -4, 230, 6, col = rgb(0.9, 0.2, 0.2, 0.15), border = NA)

# Incremental
plot(rnorm(n, mean = (1:n)/100, sd = 1), type = "l", col = "#457B9D",
     main = "Incremental Drift", xlab = "Time", ylab = "Value")

# Recurring
plot(sin((1:n)/25) * 2 + rnorm(n, 0, 0.3), type = "l", col = "#457B9D",
     main = "Recurring Drift", xlab = "Time", ylab = "Value")

par(mfrow = c(1, 1))
```

# Available Methods

| Method | Input Type | Best For |
|:-------|:-----------|:---------|
| **DDM, EDDM, HDDM-A/W** | Binary (0/1) errors | Model error monitoring |
| **KSWIN, ADWIN, Page-Hinkley** | Any numeric | Continuous data streams |
| **KL-Divergence** | Categorical | Distribution comparison |
| **ProfileDifference** | Model PDPs | Model behavior changes |

---

# Part 1: Error-Rate Methods (Binary Data)

These methods monitor **prediction errors** (0 = correct, 1 = error) from a deployed model.

## Example: Monitoring Classification Errors

```{r error-rate-setup}
library(datadriftR)
set.seed(123)

# Simulate model errors: initially 5% error rate, then degrades to 30%
n_good <- 500
n_bad <- 500
error_stream <- c(

  rbinom(n_good, 1, prob = 0.05),
  rbinom(n_bad, 1, prob = 0.30)
)
true_drift <- n_good + 1
```

### DDM (Drift Detection Method)

Monitors error rate $p$ and standard deviation $s$. Signals drift when:
$$p_t + s_t \geq p_{min} + 3 \cdot s_{min}$$

```{r ddm-example}
ddm <- DDM$new(warning_level = 2.0, out_control_level = 3.0)

warnings <- c(); drifts <- c()
for (i in seq_along(error_stream)) {
  ddm$add_element(error_stream[i])
  if (ddm$warning_detected && !ddm$change_detected) warnings <- c(warnings, i)
  if (ddm$change_detected) { drifts <- c(drifts, i); ddm$reset() }
}

cat("True drift:", true_drift, "| DDM detected:", drifts[1],
    "(delay:", drifts[1] - true_drift, ")\n")
```

### EDDM (Early Drift Detection Method)

More sensitive to gradual drift by monitoring distance between errors.

```{r eddm-example}
eddm <- EDDM$new()

drifts <- c()
for (i in seq_along(error_stream)) {
  eddm$add_element(error_stream[i])
  if (eddm$change_detected) { drifts <- c(drifts, i); eddm$reset() }
}

cat("True drift:", true_drift, "| EDDM detected:", drifts[1], "\n")
```

### HDDM (Hoeffding's Bound Methods)

Statistical guarantees via Hoeffding inequality.

```{r hddm-example}
# HDDM-A (averaging)
hddm_a <- HDDM_A$new()
drifts_a <- c()
for (i in seq_along(error_stream)) {
  hddm_a$add_element(error_stream[i])
  if (hddm_a$change_detected) { drifts_a <- c(drifts_a, i); hddm_a$reset() }
}

# HDDM-W (weighted)
hddm_w <- HDDM_W$new()
drifts_w <- c()
for (i in seq_along(error_stream)) {
  hddm_w$add_element(error_stream[i])
  if (hddm_w$change_detected) { drifts_w <- c(drifts_w, i); hddm_w$reset() }
}

cat("HDDM-A detected:", drifts_a[1], "| HDDM-W detected:", drifts_w[1], "\n")
```

### Comparing Error-Rate Methods

```{r error-rate-comparison, fig.height=4}
methods <- c("ddm", "eddm", "hddm_a", "hddm_w")
results <- sapply(methods, function(m) {
  res <- detect_drift(error_stream, method = m)
  if (nrow(res) > 0) min(res$index) else NA
})

# Visualize error rate over time
window <- 50
error_rate <- sapply(seq_along(error_stream), function(i) {
  mean(error_stream[max(1, i-window+1):i])
})

plot(error_rate, type = "l", col = "gray50", lwd = 2,
     xlab = "Observation", ylab = paste0("Error Rate (", window, "-obs window)"),
     main = "Error-Rate Method Comparison")
abline(v = true_drift, col = "black", lty = 2, lwd = 2)

colors <- c("#E63946", "#F4A261", "#2A9D8F", "#9B5DE5")
for (i in seq_along(results)) {
  if (!is.na(results[i])) abline(v = results[i], col = colors[i], lwd = 2)
}

legend("topleft", c("True drift", toupper(methods)),
       col = c("black", colors), lty = c(2, rep(1, 4)), lwd = 2, cex = 0.8)
```

---

# Part 2: Distribution Methods (Continuous Data)

These methods work with **any numeric stream**—sensor readings, feature values, predictions.

## Example: Sensor Monitoring

```{r continuous-setup}
set.seed(456)

# Temperature sensor: normal operation, then malfunction
n_normal <- 300
n_faulty <- 200
sensor_stream <- c(
  rnorm(n_normal, mean = 20, sd = 1),   # Normal: 20°C
  rnorm(n_faulty, mean = 28, sd = 2)    # Faulty: 28°C
)
true_drift <- n_normal + 1
```

### KSWIN (Kolmogorov-Smirnov Windowing)

Compares distributions using the KS test.

```{r kswin-example}
kswin <- KSWIN$new(alpha = 0.005, window_size = 100, stat_size = 30)

drifts <- c()
for (i in seq_along(sensor_stream)) {
  kswin$add_element(sensor_stream[i])
  if (kswin$detected_change()) drifts <- c(drifts, i)
}

cat("True drift:", true_drift, "| KSWIN detected:", drifts[1], "\n")
```

### Page-Hinkley Test

Monitors cumulative sum for mean shifts:
$$PH_t = \sum_{i=1}^{t}(x_i - \bar{x}_t - \delta)$$

```{r pagehinkley-example}
ph <- PageHinkley$new(threshold = 30, delta = 0.01)

drifts <- c()
for (i in seq_along(sensor_stream)) {
  ph$add_element(sensor_stream[i])
  if (ph$detected_change()) { drifts <- c(drifts, i); ph$reset() }
}

cat("True drift:", true_drift, "| Page-Hinkley detected:", drifts[1], "\n")
```

### ADWIN (Adaptive Windowing)

Automatically adjusts window size—no tuning needed.

```{r adwin-example}
adwin <- ADWIN$new(delta = 0.002)

drifts <- c()
for (i in seq_along(sensor_stream)) {
  adwin$add_element(sensor_stream[i])
  if (adwin$detected_change()) drifts <- c(drifts, i)
}

cat("True drift:", true_drift, "| ADWIN detected:", drifts[1], "\n")
```

### Comparing Distribution Methods

```{r distribution-comparison, fig.height=4}
methods <- c("kswin", "adwin", "page_hinkley")
results <- sapply(methods, function(m) {
  res <- detect_drift(sensor_stream, method = m)
  if (nrow(res) > 0) min(res$index) else NA
})

plot(sensor_stream, type = "l", col = "gray50",
     xlab = "Time", ylab = "Temperature (°C)",
     main = "Distribution Method Comparison")
abline(v = true_drift, col = "black", lty = 2, lwd = 2)

colors <- c("#E63946", "#2A9D8F", "#9B5DE5")
for (i in seq_along(results)) {
  if (!is.na(results[i])) abline(v = results[i], col = colors[i], lwd = 2)
}

legend("topleft", c("True drift", toupper(methods)),
       col = c("black", colors), lty = c(2, rep(1, 3)), lwd = 2, cex = 0.8)
```

---

# Part 3: Batch Processing with `detect_drift()`

For quick analysis, use the `detect_drift()` function:
```{r detect-drift-batch}
# Error-rate data
errors <- c(rbinom(300, 1, 0.1), rbinom(200, 1, 0.4))
detect_drift(errors, method = "ddm")

# Continuous data
temps <- c(rnorm(200, 20, 1), rnorm(200, 25, 1))
detect_drift(temps, method = "kswin", alpha = 0.01, window_size = 50)
```

---

# Part 4: ProfileDifference (Model Behavior)

Compare Partial Dependence Profiles (PDPs) to detect changes in model behavior.

```{r profile-diff, fig.height=4, message=FALSE}
# Simulate two different model behaviors
set.seed(789)
x <- seq(0, 10, length.out = 100)

# Model 1: linear relationship
profile1 <- list(x = x, y = 2 * x + rnorm(100, 0, 0.5))

# Model 2: relationship changed (drift)
profile2 <- list(x = x, y = -0.5 * x + 15 + rnorm(100, 0, 0.5))

# Calculate difference
pdiff <- ProfileDifference$new(method = "pdi")
pdiff$set_profiles(profile1, profile2)
result <- pdiff$calculate_difference()

# Visualize
par(mfrow = c(1, 2))
plot(profile1$x, profile1$y, type = "l", lwd = 2, col = "#2A9D8F",
     xlab = "Feature", ylab = "Partial Dependence", main = "PDP Comparison")
lines(profile2$x, profile2$y, lwd = 2, col = "#E63946")
legend("topleft", c("Before", "After"), col = c("#2A9D8F", "#E63946"), lwd = 2)

plot(result$x, result$der1, type = "l", lwd = 2, col = "#2A9D8F",
     xlab = "Feature", ylab = "Derivative", main = "Derivative Comparison")
lines(result$x, result$der2, lwd = 2, col = "#E63946")
legend("topright", c("Before", "After"), col = c("#2A9D8F", "#E63946"), lwd = 2)
par(mfrow = c(1, 1))

cat("Profile difference (PDI):", round(result$distance, 4), "\n")
```

---

# Quick Reference

| Scenario | Recommended Method | Code |
|:---------|:-------------------|:-----|
| Classification errors | DDM, EDDM | `detect_drift(errors, "ddm")` |
| Sensor/measurement data | KSWIN, Page-Hinkley | `detect_drift(data, "kswin")` |
| Unknown drift type | ADWIN | `detect_drift(data, "adwin")` |
| Gradual drift | EDDM, ADWIN | `detect_drift(data, "eddm")` |
| Model behavior | ProfileDifference | `ProfileDifference$new()` |

---

For detailed documentation, see the [Reference](../reference/index.html) page.

**GitHub**: [github.com/ugurdar/datadriftR](https://github.com/ugurdar/datadriftR)
